---
title: "introduction to veritas"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction to veritas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The veritas package is a set of functions developped to analyze data
generated with the Veritas questionnaire, an interactive map-based questionnaire
to collect related information on people social networks and activity space.
For more information on this questionnaire, 
see [Naud et al., 2020](https://doi.org/10.1016/j.healthplace.2020.102454).

# Input data

This package need four different tables with specific formats to work properly:

  1. `location` The set of all locations visited by the participants
  2. `people` The set of all people seen by the participants
  3. `groups` The set of all groups seen by the participants
  4. `relations` The set of all relations in-between locations, people and groups

Note that these tables can be names differently, as they will be used
as a function parameters (see following sections).

These tables need to have some specificity described below.
These can be automaticaly tested through `validateData()` function.

## Required columns

The following columns are mandatory for the package to work properly.
The column names **must** be those discussed below. The tables can contains
other variables that are useful for the user.

  - `locations`
    - `pid` The unique participant identifiers
    - `location_id` The unique location identifiers

  - `people`
    - `pid` The unique participant identifiers
    - `people_id` The unique people identifiers

  - `groups`
    - `pid` The unique participant identifiers
    - `group_id` The unique group identifiers
    - `group_size` The size of each group

  - `relations`
    - `pid` The unique participant identifiers
    - `node_1` The unique identifiers of the first node
    - `node_2` The unique identifiers of the second node
    - `relation_type` The relation type 
    (see below for further detail on this variable)

## Locations, people and groups identifiers
All `locations_id` and `people_id` and `group_id` must be unique.

## Participant identifiers

All participant identifiers (i.e., `pid`) found in `people`, `groups`, 
and `relations` tables must be found in the `locations` table.

According to Veritas data collection protocol, should identify at least
their home as an activity locations, and therefore must be found in the
`locations` table.

## Relation type

`relation_type` is a categorical variables that can take three numerical
values:

  - `1` The relation in-between people (i.e., who know whom)
  - `2` The relation between people and groups 
  (i.e., who is part of which groups)
  - `3` The relation between location and people / group 
  (i.e., In which locations people or groups are seen).

## node_1 and node_2

In the `relation` table, `node_1` and `node_2` identify specific elements
depending on the `relation_type`:

  - `relation_type == 1`
    - `node_1` and `node_2` are `people_id`

  - `relation_type == 2`
    - `node_1` are `people_id`
    - `node_2` are `group_id`
  - `relation_type == 3`
    - `node_1` are `location_id`
    - `node_2` are `people_id` or `group_id`

## Group size

All group size should be greater than 0, and this when removing for the 
people that belong to groups (i.e., `relation_type == 2`)


# Load data

`locations`, `people`, `groups` and `relations` table are loaded using the 
`transformData()` function.

```{r load, message = FALSE}
# Load library
library(veritas)

# Load veritas data
veritas_data <- transformData(veritas::locations,
                              veritas::people,
                              veritas::groups,
                              veritas::relations)
```

# Validation

`transformData()` check if input datasets are valid by calling two functions, 
`validateData()` and `validateNAs()`. These functions can be called directly
to obtain a more detailed assessment.

```{r validate data}
# Load veritas data
validateData(veritas::locations,
             veritas::people,
             veritas::groups,
             veritas::relations)
```

```{r validate NAs}
# Load veritas data
validateNAs(veritas::locations,
             veritas::people,
             veritas::groups,
             veritas::relations)
```

# Social network analysis

Veritas data can be transformed into social networks using
`relation_type %in% c(1, 2)`.

## Generate social networks
First we need to create a social network object

```{r create social networks, eval = FALSE}
# Create social network
social_networks <- createSocialNetworks(veritas_data)
```

## Calculate measures
Then we can calculate social network indicators

```{r calculate social measures, eval = FALSE}

# Calculate base social network indicator
base_sn_measures <- socialNetworkMeasures(social_networks)

# Calcul SN measures specific to InterAct project
# THIS WONT WORK WITH DUMMY DATA SETS!!
interact_sn_measures <- interactSnMeasures(veritas_data)

# Combine both data frames
sn_measures <- merge(base_sn_measures, interact_sn_measures, 
                     by = "pid")

```

